Efficient ancestry and mutation simulation with msprime 1.0

### 2022-01-06 paper.tex
error in description of mutation algorithm

Since the general algorithm above applies mutations at a single rate
independent of ancestral state,
a model in which different alleles mutate at different rates
must necessarily produce some silent mutations,
i.e., mutations in which the derived allele is equal to the parental allele.
To illustrate this, consider a mutation model
in which $A$ or $T$ mutates to a randomly chosen different nucleotide at rate $\alpha$
and $C$ or $G$ mutates at rate $\beta$, with $\beta < \alpha$.
To implement this, first place mutations at the largest total rate, which is $\alpha$.
Then, at each site, choose an ancestral allele from the root distribution,
and for each mutation, choose a derived allele as follows:
if the parental allele is $A$ or $T$, then choose a random derived allele
different to the parental allele;
if the parental allele is $C$ or $G$,
then choose the derived allele to be equal to the parent allele
with probability $\beta/(\alpha + \beta)$,
and randomly choose a different nucleotide otherwise.
This produces the correct distribution by Poisson thinning:
a Poisson process with rate $\alpha$ in which each point is discarded independently
with probability $\beta / (\alpha + \beta)$ is equivalent to a Poisson
process with rate $\beta$.
All finite-state models (implemented under the generic \texttt{MatrixMutationModel} class)
work in this way: mutations are placed at the maximum mutation rate,
and then some silent mutations will result.

>>>>

Since the general algorithm above applies mutations at a single rate
independent of ancestral state,
a model in which different alleles mutate at different rates
must necessarily produce some silent mutations,
i.e., mutations in which the derived allele is equal to the parental allele.
To illustrate this, consider a mutation model
in which $A$ or $T$ mutates to a randomly chosen different nucleotide at rate $\alpha$
and $C$ or $G$ mutates at rate $\beta$, with $\beta < \alpha$.
To implement this, first place mutations at the largest total rate, which is $\alpha$.
Then, at each site, choose an ancestral allele from the root distribution,
and for each mutation, choose a derived allele as follows:
if the parental allele is $A$ or $T$, then choose a random derived allele
different to the parental allele;
if the parental allele is $C$ or $G$,
then choose the derived allele to be equal to the parent allele
with probability $1-\beta/\alpha$,
and randomly choose a different nucleotide otherwise.
This produces the correct distribution by Poisson thinning:
a Poisson process with rate $\alpha$ in which each point is kept independently
with probability $\beta / \alpha$ is equivalent to a Poisson
process with rate $\beta$.
All finite-state models (implemented under the generic \texttt{MatrixMutationModel} class)
work in this way: mutations are placed at the maximum mutation rate,
and then some silent mutations will result.
